---
title: "DATA 621 Assignment 3"
author: "Critical Thinking Group 2: George Cruz Deschamps, Karim Hammoud, Maliat Islam, Matthew Lucich, Gabriella Maritnez, Ken Popkin"
date: "Date: `r Sys.Date()` | Due: 2021-11-07" 
output:
      html_document:
        toc: yes
        toc_float: yes
        theme: yeti
        highlight: kate
        font-family: "Arial"
        code_folding: "hide"
---
# Binary Logistic Regression {.tabset}
## Overview 
In this homework assignment, you will explore, analyze and model a data set containing information on crime
for various neighborhoods of a major city. Each record has a response variable indicating whether or not the crime
rate is above the median crime rate (1) or not (0).  
  
Your objective is to build three different binary logistic regression models on the training data set to predict whether 
or not neighborhoods are at risk for high crime. You will provide classifications and probabilities for the
evaluation data set using your binary logistic regression model. You can only use the variables given to you (or
variables that you derive from the variables provided). Below is a short description of the variables of interest in
the data set:  
  
* `zn`       proportion of residential land zoned for lots over 25,000 sq.ft.
* `indus`    proportion of non-retail business acres per town
* `chas`     Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)
* `nox`      nitrogen oxides concentration (parts per 10 million)
* `rm`       average number of rooms per dwelling
* `age`      proportion of owner-occupied units built prior to 1940
* `dis`      weighted distances to five Boston employment centers
* `rad`      index of accessibility to radial highways
* `tax`      full-value property-tax rate per $10,000
* `ptratio`  pupil-teacher ratio by town
* `lstat`    % lower status of the population
* `medv`     median value of owner-occupied homes in $1000's
* `target`   indicating whether or not the crime rate is above the median crime rate (1) or not (0) **response variable** 

## Package Dependencies

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(Hmisc)
library(psych)
library(corrplot)
library(RColorBrewer)
library(knitr)
library(MASS)
library(caret)
library(kableExtra)
library(ResourceSelection)
library(pROC)
library(Amelia)
library(summarytools)
```

# Data Exploration {.tabset}
```{r}
#load data
crime_train <- read.csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Assignment_3/data/crime-training-data_modified.csv")
crime_test <- read.csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Assignment_3/data/crime-evaluation-data_modified.csv")
```

Using dplyr's ```glimpse()```^[https://www.rdocumentation.org/packages/dplyr/versions/0.3/topics/glimpse] function, we can take a 'glimpse" into the data and easily see the dimensions, variable names and types.

```{r}
glimpse(crime_train)
 
```

```{r}
descr(crime_train,
  headings = FALSE, #remove headings
  stats = "common",# most common descriptive statistics, default is all
  transpose = TRUE #allows for better display due to large amount of variables
  ) %>% 
  kbl(caption = "Univariate Descriptive Statistics - Training Data Set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

```{r}
missmap(crime_train, main = "Missing vs Observed Values")
```

In the histogram plot below, we see that `medv`, and `rm` are normally distributed. We also see bi-modal distribution of the variables `indus`, `rad` and `tax`. The rest of the variables show moderate to high skewness on either side respectively.
```{r}
# distribution of the varaibles
crime_train %>%
  gather(key, value, -c(target, chas)) %>%
  ggplot(aes(value)) +
  geom_histogram(binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3)), fill="steelblue") +
  facet_wrap(~ key, scales = 'free', ncol = 3) +
  theme_minimal()
```

In the box-plot figure below, we see many variables exhibit outliers We also see very high interquartile range for rad and tax variables where crime rate is above the median. Lastly, the variance between the 2 values of target differs for `zn`, `nox`, `age`, `dis`, `rad` & `tax`, which indicates that we will want to consider adding quadratic terms for them.

```{r}
crime_train %>%
  dplyr::select(-chas) %>% 
  gather(key, value, -target) %>% 
  mutate(key = factor(key),
         target = factor(target)) %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot(aes(fill = target)) +
  facet_wrap(~ key, scales = 'free', ncol = 3) +
  scale_fill_manual(values=c("steelblue", "firebrick")) +
  theme_minimal()
```

In the correlation table and plot below, we see moderate positive correlation between variables nox, age, rad, tax, indus and target variables; and moderate negative correlation between variable dis. And the rest of the variables have weak or no correlations.
```{r}
kable(sort(cor(dplyr::select(crime_train, target, everything()))[,1], decreasing = T), col.names = c("Correlation")) %>% 
  kable_styling(full_width = F)
```


```{r fig.height = 7, fig.width = 7, warning=FALSE}
numeric_values <- crime_train %>% select_if(is.numeric)
train_cor <- cor(numeric_values)
corrplot.mixed(train_cor, tl.col = 'black', tl.pos = 'lt', upper = "number", lower="circle")
```

# Data Preparation {.tabset}

Some of the variables are skewed, have outliers or follow a bi-modal distribution. Therefore, we will perform transformation on some of these variables. First, we will remove the variable `tax` due to multicollinearity and it’s high VIF score. Next, we will take log() transformation of `age` and `lstat` variables to lower skewness. Lastly, we will add quadratic term to `zn`, `rad`, and `nox` variables to account for its variances with respect to target variable.

```{r}
# find multicollinear variables
kable((car::vif(glm(target ~. , data = crime_train))), col.names = c("VIF Score")) %>%  #remove tax for high vif score
  kable_styling(full_width = F)
```

```{r}
# transformation of the variables. 
crime_train_trans <- crime_train %>%
  dplyr::select(-tax) %>% 
  mutate(age = log(age),
         lstat = log(lstat),
         zn = zn^2,
         rad = rad^2,
         nox = I(nox^2))
```

```{r}
# histogram distribution of the transformed variables
crime_train_trans%>%
  gather(key, value, -c(target, chas)) %>%
  ggplot(aes(value)) +
  geom_histogram(binwidth = function(x) 2 * IQR(x) / (length(x)^(1/3)), fill="steelblue") +
  facet_wrap(~ key, scales = 'free', ncol = 3) +
  theme_minimal()
```


# Build Models {.tabset}

# Select Models {.tabset}

