---
title: "DATA 621 Assignment 3"
author: "Critical Thinking Group 2: George Cruz Deschamps, Karim Hammoud, Maliat Islam, Matthew Lucich, Gabriella Maritnez, Ken Popkin"
date: "Date: `r Sys.Date()` | Due: 2021-11-07" 
output:
      html_document:
        toc: yes
        toc_float: yes
        theme: yeti
        highlight: kate
        font-family: "Arial"
        code_folding: "show"
---
# Binary Logistic Regression {.tabset}
## Overview 
In this homework assignment, you will explore, analyze and model a data set containing information on crime
for various neighborhoods of a major city. Each record has a response variable indicating whether or not the crime
rate is above the median crime rate (1) or not (0).  
  
Your objective is to build three different binary logistic regression models on the training data set to predict whether 
or not neighborhoods are at risk for high crime. You will provide classifications and probabilities for the
evaluation data set using your binary logistic regression model. You can only use the variables given to you (or
variables that you derive from the variables provided). Below is a short description of the variables of interest in
the data set:  
  
* `zn`       proportion of residential land zoned for lots over 25,000 sq.ft.
* `indus`    proportion of non-retail business acres per town
* `chas`     Charles River dummy variable (= 1 if tract bounds river; 0 otherwise)
* `nox`      nitrogen oxides concentration (parts per 10 million)
* `rm`       average number of rooms per dwelling
* `age`      proportion of owner-occupied units built prior to 1940
* `dis`      weighted distances to five Boston employment centers
* `rad`      index of accessibility to radial highways
* `tax`      full-value property-tax rate per $10,000
* `ptratio`  pupil-teacher ratio by town
* `lstat`    % lower status of the population
* `medv`     median value of owner-occupied homes in $1000's
* `target`   indicating whether or not the crime rate is above the median crime rate (1) or not (0) **response variable** 

## Package Dependencies

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(Hmisc)
library(psych)
library(corrplot)
library(RColorBrewer)
library(knitr)
library(MASS)
library(caret)
library(kableExtra)
library(ResourceSelection)
library(pROC)
library(Amelia)
library(summarytools)
library(DataExplorer)
library(knitr)
library(dplyr)
```

# Data Exploration {.tabset}
```{r}
#load data
crime_train <- read.csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Assignment_3/data/crime-training-data_modified.csv")
crime_test <- read.csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Assignment_3/data/crime-evaluation-data_modified.csv")
```


Exploratory data analysis is the process to get to know your data, so that a hypothesis can be generated and later tested. Visualization techniques are usually applied to aid the exploration of the data.

To get introduced to the dataset, we use DataExplorer's `introduce()` function:

```{r}
kable(t(introduce(crime_train)), row.names = TRUE, col.names = "", format.args = list(big.mark = ","))
```
  
  
Using dplyr's ```glimpse()```^[https://www.rdocumentation.org/packages/dplyr/versions/0.3/topics/glimpse] function, we can take a "glimpse" into the both the `crime_train` and `crime_test` data respectively, and easily see the dimensions, variable names and types.  
  
From the `glimpse()`, in the `crime_train` dataset, we confirm the two variables `chas` and `target` are factors^[https://www.stat.berkeley.edu/~s133/factors.html#] as noted in the description of variables above. These variables will be transformed in our data preparation stage.
```{r}
glimpse(crime_train)
 
glimpse(crime_test)
```
  

Next, we proceed with our exploratory data analysis by providing univariate descriptive statistics on our training dataset, `crime_train`.

```{r}
descr(crime_train,
  headings = FALSE, #remove headings# 
  transpose = TRUE #allows for better display due to large amount of variables
  ) %>% 
  kbl(caption = "Univariate Descriptive Statistics - Training Data Set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```


Furthermore, in the histogram plot below, we see that `medv`, and `rm` are normally distributed. We also note bi-modal distribution of the variables `indus`, `rad` and `tax`. The rest of the variables show moderate to high skewness on either side respectively.
```{r}
# distribution of the varaibles
crime_train %>% 
  gather(variable, value, zn:target) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "steelblue", color="steelblue") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

In the box-plot figure below, we see many variables exhibit outliers We also see very high interquartile range for `rad` and `tax` variables where crime rate is above the median. Lastly, the variance between the 2 values of target differs for `zn`, `nox`, `age`, `dis`, `rad` & `tax`, which indicates that we will want to consider adding quadratic terms for them.

```{r}
crime_train %>%
  dplyr::select(-chas) %>% 
  gather(key, value, -target) %>% 
  mutate(key = factor(key),
         target = factor(target)) %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot(aes(fill = target)) +
  facet_wrap(~ key, scales = 'free', ncol = 3) +
  scale_fill_manual(values=c("steelblue", "firebrick")) +
  theme_minimal()
```



In order to investigate if there is existing correlation between the data and the target variable, we obtain the values of correlation as well as a visualization. 

The correlation table and plot below, we see moderate positive correlation between variables `nox`, `age`, `rad`, `tax`, `indus` and `target` variables; and moderate negative correlation between variable dis. And the rest of the variables have weak or no correlations.
```{r}
kable(sort(cor(dplyr::select(crime_train, target, everything()))[,1], decreasing = T), col.names = c("Correlation")) %>% 
  kable_styling(full_width = F)

```
  
Below is a correlation matrix of the feature variables in our dataset. The correlation matrix confirms that multicollinearity is a concern.

```{r fig.height = 7, fig.width = 7, warning=FALSE}
numeric_values <- crime_train %>% select_if(is.numeric)
train_cor <- cor(numeric_values)
corrplot.mixed(train_cor, tl.col = 'black', tl.pos = 'lt', upper = "number", lower="circle")
```
  
Lastly, we proceed to check if there are any missing data points in the `crime_train`

```{r}
map(crime_train, ~sum(is.na(.))) %>% t()
```

```{r}
missmap(crime_train, main = "Missing vs Observed Values")
```



# Data Preparation {.tabset}
**Feature Engineering**  
In an effort to fine tune our model, we will introduce the use of feature engineering on select variables.  

* `ptratio_indicator` assigned a value of 1 if the pupil to teacher ratio is > 16, 0 if `ptratio` is greater than 16 ^[https://www.publicschoolreview.com/average-student-teacher-ratio-stats/national-data]  
* `lstat_indicator` assigned a value of 1 if > 15 % of the population is considered low status, 0 otherwise
* `dis_indicator` assigned a value of 1 if the distance from employment centers is > 4, 0 if `dis` is less than 4 (mean value: 3.8)

```{r}
#make a copy of original dataset
train <- crime_train

#convert chas and target to factors
train$chas <- as.factor(train$chas)
train$target <- as.factor(train$target)

#add new variables
train$ptratio_indicator <- as.factor(ifelse(train$ptratio < 16, 1, 0))
train$lstat_indicator <- as.factor(ifelse(train$lstat < 15, 1, 0))
train$dis_indicator <- as.factor(ifelse(train$dis < 4, 1, 0))
train$age_greater_than_77 <- as.factor(ifelse(train$age >= 77, 1, 0)) #median age is 77
```


```{r}
str(train)
```



**Data Transformation**
Some of the variables are skewed, have outliers or follow a bi-modal distribution. Therefore, we will perform transformation on some of these variables. First, we will remove the variable `tax` due to multicollinearity and it’s high VIF score. Next, we will take log() transformation of `age` and `lstat` variables to lower skewness. Lastly, we will add quadratic term to `zn`, `rad`, and `nox` variables to account for its variances with respect to target variable.

```{r}
#MI find multicollinear variables
kable((car::vif(glm(target ~. , data = crime_train))), col.names = c("VIF Score")) %>%  #remove tax for high vif score
  kable_styling(full_width = F)
```

```{r}
# MI transformation of the variables. 
trans_train <- train %>%
  dplyr::select(-tax) %>% 
  mutate(age = log(age),
         lstat = log(lstat),
         zn = zn^2,
         rad = rad^2,
         nox = I(nox^2))
```

```{r message=FALSE, warning=FALSE}
# MI histogram distribution of the transformed variables
trans_train %>% 
  gather(key, value, c(age, lstat, zn, rad, nox)) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "steelblue", color="steelblue") + 
  facet_wrap(~ key, scales = 'free', ncol = 3) +
  labs(x = element_blank(), y = element_blank())

```

```{r}
str(trans_train)
```



# Build Models {.tabset}

First, we begin by building the null binary regression model. We will use this model to compare it to the subsequent models we build.

Coefficients for the null or Intercept only model:
```{r}
null_model <- glm(target~1,data=train, family="binomial"(link = "logit"))
summary(null_model)
```

Next we proceed to make our first model using the both the original, untransformed variables and the engineered features.
```{r}
model_1 <- glm(target~., family = "binomial"(link = "logit"), data = train)

summary(model_1)
```
  
Confidence intervals for the regression coefficients:
```{r}
confint.default(model_1)

exp(cbind(OR=coef(model_1), confint.default(model_1)))
```
  
The odds ratio would indicate the multiplicative change in odds of crime for every one unit increase on a predictor variable.  
Odds-ratios for coefficients:
```{r}
exp(coef(model_1))
```
    
For our second model, we will use the `trans_train` dataset which includes the transformed variables in the previous section in addition to the engineered features.
```{r}
model_2 <- glm(target~ .,family = binomial(link = "logit"), data = trans_train)

summary(model_2)
```

For our third model: 
```{r}
trans_train_mod_3 <- trans_train %>% 
  dplyr::select(1:12)

model_3 <- glm(target ~ . -rm -chas - age -indus, 
             family = binomial(link = "logit"), 
             trans_train_mod_3)
summary(model_3)
```


# Select Models {.tabset}
To determine if the models produced have significant improvement in fit over the null model, we make use of the `anova()` function.
```{r}
anova(null_model,model_1,model_2,model_3,test="LRT")
```
