---
title: "FinalFinal1"
author: "Maliat I"
date: "12/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## R Markdown

### Overview
There are millions of stray pets around the world, some of which are fortunate enough to be adopted while many others are not.  While adoption of a pet is often the definition of success, the rate at which a pet is adopted is also a key success factor - pets that take a long time to adopt contribute to over-crowded animal shelters and can prevent taking on new strays.  Sadly, pets that are not adopted eventually need to be euthanized.

### Learn more about the data
[About the data](https://www.kaggle.com/c/petfinder-adoption-prediction/data)

### What to Predict
Predictor (Adoption Speed) Description: Predict how quickly, if at all, a pet is adopted.

The values are determined in the following way:
0 - Pet was adopted on the same day as it was listed.
1 - Pet was adopted between 1 and 7 days (1st week) after being listed.
2 - Pet was adopted between 8 and 30 days (1st month) after being listed.
3 - Pet was adopted between 31 and 90 days (2nd & 3rd month) after being listed.
4 - No adoption after 100 days of being listed.

### This Notebook...
The data has no missing values, but there are a number of features in text that need to be converted to some numeric value.  This notebook performs those changes.

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(dplyr)
library(reshape)
library(ggplot2)
library(purrr)
library(psych)
library(tidyr)
```

```{r}
library(scales)
```


### Load Data
Load the data
```{r echo=FALSE}
# load data
data <- read.csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Final%20Project/data/TrainingData/train.csv")
#data <- read.csv("C:\\Users\\malia\\OneDrive\\Desktop\\MSDSDATA621\\Data_621\\Final Project\\data\\TestData\\test.csv\")
head(data,2)
```
### Missing Values Count 
There are no missing values
```{r}
map(data, ~sum(is.na(.))) %>% t()
```


### Show the type of each data
```{r}
str(data)
```



### Adoption rate
### It can be observed that the rate of not getting adopted after 100 days of being listed is more frequent. 

```{r}

 data %>%
  ggplot(aes(x= AdoptionSpeed, fill = AdoptionSpeed)) +
  geom_bar(stat = "count", color = "black") +
  theme_minimal() +
  theme(axis.title.y = element_blank()) +
  scale_y_continuous(labels = comma) +
  scale_fill_brewer(palette="blue") +
  theme(legend.position = "top") 
```
### Exploration
```{r}
library(tidyverse)
library(jsonlite)
library(scales)
library(lubridate)
library(repr)
library(ggrepel)
library(gridExtra)
library(tidytext)
library(grid)
library(rjson)
library(xgboost)
library(caret)
library(Metrics)
library(Ckmeans.1d.dp)


```

```{r}
library(dplyr)
train <- read_csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Final%20Project/data/TrainingData/train.csv")
test <- read_csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Final%20Project/data/TestData/test.csv")
state_labels <- read_csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Final%20Project/data/state_labels.csv")
breed_labels <- read_csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Final%20Project/data/TrainingData/breed_labels.csv")
color_labels <- read_csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Final%20Project/data/color_labels.csv")
tr_te <- bind_rows(train, test)


train <- left_join(train, breed_labels %>%dplyr:: select(Breed1=BreedID, MainBreed=BreedName), by="Breed1")
train <- left_join(train, breed_labels %>%dplyr:: select(Breed2=BreedID, SecondBreed=BreedName), by="Breed2")
train <- left_join(train, color_labels %>%dplyr:: select(Color1=ColorID, ColorName1=ColorName), by="Color1")
train <- left_join(train, color_labels %>% dplyr::select(Color2=ColorID, ColorName2=ColorName), by="Color2")
train <- left_join(train, color_labels %>% dplyr::select(Color3=ColorID, ColorName3=ColorName), by="Color3")

train <- train %>% dplyr::select(-State, -Breed1, -Breed2, - Color1, -Color2, -Color3)

train <- train %>% mutate_at(vars(Type, Gender, AdoptionSpeed), as.factor)
train <- train %>% mutate(Type=recode(Type, "1"= "Dog", "2"= "Cat"),
                         Gender=recode(Gender, "1"= "Male", "2" = "Female", "3"= "Mixed"),
                         AdoptionSpeed=recode(AdoptionSpeed,
                                              "0"= "0 - Adopted on the same day",
                                              "1" = "1 - Adopted between 1 and 7 days",
                                              "2" = "2 - Adopted between 8 and 30 days",
                                             "3" = "3 - Adopted between 31 and 90 days",
                                            "4" = "4 - No adoption after 100 days"))
train <- train %>% mutate_if(is_character, as_factor)
glimpse(train)
```
### Pure Breed Feature:
### We will observe if pure breed pets are getting adopted faster than the pets that are not pure breed.


```{r}
not_pure <- c("Domestic Short Hair", "Domestic Medium Hair", "Domestic Long Hair", "Mixed Breed")
train$pure_breed <- ifelse(train$MainBreed %in% not_pure, 0, 1)

train %>% filter(pure_breed==1) %>% count(Type, MainBreed) %>% group_by(Type) %>% top_n(10, n) %>%
ggplot(aes(x=reorder(MainBreed, -n), y=n))+
geom_bar(stat="identity", fill="blue") +
facet_wrap(~Type, scales = "free_x") +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
labs(x="Most common pure breeds", y="number of pets")
```
```{r}
train %>% count(AdoptionSpeed, pure_breed) %>%
ggplot(aes(x=AdoptionSpeed, y=n, fill=as.factor(pure_breed))) +
geom_bar(stat="identity", position="fill") +
theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
guides(fill=guide_legend(title="Pure Breed")) +
scale_y_continuous(labels=percent) +
labs(x="", y="percent")
```
### As pure breed pets have faster adoption rate, a new variable is created to identify pure breed pets.

```{r}
tr_te <- left_join(tr_te, state_labels %>% dplyr::rename(State=StateID), by="State")
tr_te <- left_join(tr_te, breed_labels %>% dplyr::select(Breed1=BreedID, MainBreed=BreedName), by="Breed1")

#creating Has Name variable
tr_te$has_name <- ifelse(is.na(tr_te$Name), 0, 1)

#creating Pure Breed variable
not_pure <- c("Domestic Short Hair", "Domestic Medium Hair", "Domestic Long Hair", "Mixed Breed")
tr_te$pure_breed <- ifelse(tr_te$MainBreed %in% not_pure, 0, 1)

#making Not Specified in ordinal factors NA (just in case there are any in stage 2)
tr_te$MaturitySize[tr_te$MaturitySize==0] <- NA
tr_te$FurLength[tr_te$FurLength==0] <- NA
tr_te$Health[tr_te$Health==0] <- NA

categorical_vars <- c("Type", "Gender", "Vaccinated", "Dewormed", "Sterilized", "StateName", "MainBreed", "has_name", "pure_breed", "Breed2", "Color1", "Color2", "Color3")

tr_te <- tr_te %>% dplyr::select(-Name, -Breed1, -RescuerID, -Description, -State, -PetID) %>%
mutate_at(categorical_vars, funs(factor(.))) %>% mutate_if(is.numeric, as.integer)

glimpse(tr_te)
```

```{r}
glimpse(train)
```
```{r}
glimpse(tr_te)
```
#### Fast adoption:
### Adoption speed less than 3 is categorized as high(1) and more than 3 is categorized as low(0).
```{r}
High = ifelse(tr_te$AdoptionSpeed<3, "1", "0")
tr_te = data.frame(tr_te, High)
```

```{r}
glimpse(tr_te)
```

```{r}
tr_te$High<-as.numeric(tr_te$High)
```
```{r}
glimpse(tr_te)
```
```{r}
library(MASS)
modelnegb <- glm.nb(High ~ ., data = tr_te)

```
### Negative Binomial Model:
```{r}
summary(modelnegb)
```
### Variables with high amount of factors are dropped.
```{r include=FALSE}
newtr_te<-tr_te[c(-3,-6,-7,-16,-18,-19)]
```

### Dividing into train and test set
### Data is now partitioned into 70% train and 30% test.
```{r}
set.seed(100) 

index = sample(1:nrow(newtr_te), 0.7*nrow(newtr_te)) 

train1 = newtr_te[index,] # Create the training data 
test1 = newtr_te[-index,] # Create the test data

dim(train1)
dim(test1)
```
### Negative binomial model:
```{r}
modelnegb1 <- glm.nb(High ~ ., data = train1)
```

```{r}
summary(modelnegb1)
```
### Negative binomial model with significant variables:
```{r}
modelnegb2 <- glm.nb(High~ Age  + FurLength  + Health+
                as.factor(Gender) +
                as.factor(Vaccinated) + 
                as.factor(Sterilized)+
                as.factor(Color1), data = train1)

summary(modelnegb2)
```

```{r}
modelnegb3 <- glm.nb(High~ Age  + FurLength  + Health+
                as.factor(Gender) +
                as.factor(Color1), data = train1)

summary(modelnegb3)
```
### Linear model with significant variables:
```{r}
modellin <- lm(High~ Age  + FurLength  + Health+
                as.factor(Gender) +
                as.factor(Vaccinated) + 
                as.factor(Sterilized)+
                as.factor(Color1), data = train1)

summary(modellin)
```





```{r}
library(kableExtra)
aic1 <- modelnegb$aic
aic2 <- modelnegb1$aic
aic3 <- modelnegb2$aic 
aic4 <- modellin$aic
mse1 <- mean((train1$High - predict(modelnegb))^2)
mse2 <- mean((train1$High - predict(modelnegb1))^2)
mse3 <- mean((train1$High - predict(modelnegb2))^2)
mse4 <- mean((train1$High - predict(modellin))^2)
compare_aic_mse <- matrix(c(mse1, mse2, mse3 ,mse4 ,aic1, aic2,aic3,aic4),nrow=4,ncol=2,byrow=TRUE)


rownames(compare_aic_mse) <- c("Model1","Model2","Model3","Model4")
colnames(compare_aic_mse) <- c("MSE","AIC")
compare_models <- as.data.frame(compare_models)

kable(compare_aic_mse)  %>% 
  kable_styling(full_width = T)
```
### The negative binomial model with significant variables and lower AIC score is selected:
### We are going to deploy this model on our test data.







```{r}
pred <- predict(modelnegb2, train1, type = "response")
summary(pred)
```

```{r}
pred1 <- predict(modellin, train1, type = "response")
summary(pred1)
```

### It can be observed that based on the prediction that Mean is 0.4953 and Median is 0.518. Wemay need a different model like Binomial, Random Forest or XGBoost.