---
title: "DATA 621 Final Project"
author: "Group 2"
---

### Overview
There are millions of stray pets around the world, some of which are fortunate enough to be adopted while many others are not.  While adoption of a pet is often the definition of success, the rate at which a pet is adopted is also a key success factor - pets that take a long time to adopt contribute to over-crowded animal shelters and can prevent taking on new strays.  Sadly, pets that are not adopted eventually need to be euthanized.

### Learn more about the data
^[About the data](https://www.kaggle.com/c/petfinder-adoption-prediction/data)

### What to Predict
Predictor (Adoption Speed) Description: Predict how quickly, if at all, a pet is adopted.

The values are determined in the following way:
0 - Pet was adopted on the same day as it was listed.
1 - Pet was adopted between 1 and 7 days (1st week) after being listed.
2 - Pet was adopted between 8 and 30 days (1st month) after being listed.
3 - Pet was adopted between 31 and 90 days (2nd & 3rd month) after being listed.
4 - No adoption after 100 days of being listed.

### This notebook...
inputs the transformation training data and completes an analysis of the features, with visuals.

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(dplyr)
library(reshape)
library(ggplot2)
library(purrr)
library(psych)
library(tidyr)
library(corrplot)
```

### Load Data
Load the data
```{r echo=FALSE}
# load data
data <- read.csv('./data/TrainingData/train_cleaned.csv')
head(data,2)
```

### Correlation of each feature to AdoptionSpeed
```{r}
#Function to calculate each feature's correlation to AdoptionSpeed
correlations <- function(data) {
  res_vector  <- vector()
  colnames_vector <- colnames(data)
  
  for(i in 1:ncol(data)) {
    
    res <- cor.test(data[ , i], data$AdoptionSpeed, 
                      method = "pearson")
    
    res_round = abs(round(res$estimate,4))
    
    res_vector <- c(res_vector, res_round)
  }
  
  corrdf <- data.frame(colnames_vector, res_vector)
  corrdf_sort <- corrdf %>% arrange(desc(res_vector))
  
  #Remove AdoptionSpeed from correlation df
  corrdf_sort = corrdf_sort[-c(1),]
  
  return(corrdf_sort)
}
```

```{r}
#Dogs and Cats: each feature's correlation to AdoptionSpeed
corrdf_sort = correlations(data)

ggplot(corrdf_sort, aes(x=reorder(colnames_vector,res_vector), y=res_vector)) +
  geom_bar(stat="identity") + ggtitle("Correlations to Adoption Speed - Dogs and Cats") +
  theme(axis.text.x = element_text(angle = 90, size = 10))
```

```{r}
#Dogs: each feature's correlation to AdoptionSpeed
dog_data = filter(data, Type == 1)
dog_data = subset(dog_data, select = -c(Type))
corrdf_sort = correlations(dog_data)

#Dogs: each feature's correlation to AdoptionSpeed
ggplot(corrdf_sort, aes(x=reorder(colnames_vector,res_vector), y=res_vector)) +
  geom_bar(stat="identity") + ggtitle("Correlations to Adoption Speed - Dogs") +
  theme(axis.text.x = element_text(angle = 90, size = 10))
```

```{r}
#Cats: each feature's correlation to AdoptionSpeed
cat_data = filter(data, Type == 2)
cat_data = subset(cat_data, select = -c(Type))
corrdf_sort = correlations(cat_data)

#Dogs: each feature's correlation to AdoptionSpeed
ggplot(corrdf_sort, aes(x=reorder(colnames_vector,res_vector), y=res_vector)) +
  geom_bar(stat="identity") + ggtitle("Correlations to Adoption Speed - Cats") +
  theme(axis.text.x = element_text(angle = 90, size = 10))
```

### Visual Correlation between features
```{r}
corrplot.mixed(cor(data %>% keep(is.numeric)), tl.col = 'black', tl.pos = 'lt', upper = "number", lower="shade", shade.col=NA, tl.srt=45)
```

### Boxplot of each feature against AdoptionSpeed
```{r}
column_names = names(data)
for(i in 1:ncol(data)) {
  #ggplot(data, aes(data[ , i], data$AdoptionSpeed)) + geom_bar()
  boxplot(AdoptionSpeed ~ data[ , i], data = data, main=column_names[i])
}
```

### Barplots of each feature's value counts
```{r}
for(i in 1:ncol(data)) {
  counts <- table(data[ , i])
  barplot(counts, main=column_names[i])
}
```

