---
title: "DATA 621 Final Project"
author: "Group 2"
---

### Overview
There are millions of stray pets around the world, some of which are fortunate enough to be adopted while many others are not.  While adoption of a pet is often the definition of success, the rate at which a pet is adopted is also a key success factor - pets that take a long time to adopt contribute to over-crowded animal shelters and can prevent taking on new strays.  Sadly, pets that are not adopted eventually need to be euthanized.

### Learn more about the data
[About the data](https://www.kaggle.com/c/petfinder-adoption-prediction/data)

### What to Predict
Predictor (Adoption Speed) Description: Predict how quickly, if at all, a pet is adopted.

The values are determined in the following way:
0 - Pet was adopted on the same day as it was listed.
1 - Pet was adopted between 1 and 7 days (1st week) after being listed.
2 - Pet was adopted between 8 and 30 days (1st month) after being listed.
3 - Pet was adopted between 31 and 90 days (2nd & 3rd month) after being listed.
4 - No adoption after 100 days of being listed.

### This Notebook...
The data has no missing values, but there are a number of features in text that need to be converted to some numeric value.  This notebook performs those changes.

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(dplyr)
library(reshape)
library(ggplot2)
library(purrr)
library(psych)
library(tidyr)
```

### Load Data
Load the data
```{r echo=FALSE}
# load data
data <- read.csv('./data/TrainingData/train.csv')
#data <- read.csv('./data/TestData/test.csv')
head(data,2)
```

### Missing Values Count 
There are no missing values
```{r}
map(data, ~sum(is.na(.))) %>% t()
```


### Show the type of each data
```{r}
str(data)
```

### New features
AgeYears - provided in months this is too many variations.  Decided to convert to new feature of AgeYears and drop Age
```{r}
data$AgeYears = round(data$Age/12, 0)
```

NumColors - derived by counting the number of colors a pet has using the Color1,2, and 3 features
- keep these features, but given there are 7 color categories the value of this feature is questionable in doubt
- create a NumColors feature with a 1,2, or 3 respectively
```{r}
num_colors_vec = c()
for (i in 1:nrow(data)){
  ncolors = 1
  
  color2 <- data[i,8]
  color3 <- data[i,9]
  
  if (color2 > 0){ncolors = ncolors + 1}
  if (color3 > 0){ncolors = ncolors + 1}
  
  num_colors_vec <- c(num_colors_vec, ncolors)
}

data$NumColors = num_colors_vec
```

AllMeds = the sum of Vaccinated, Dewormed, and Sterilized.  Combining these into one feature could potentially reduce
multi-collinearity
```{r}
data$AllMeds = data$Vaccinated + data$Dewormed + data$Sterilized
```

### Remove features 
Remove the following features...
Name - this is a text field and new owners can (and usually do) rename their pets, so removing this feature
Age - replaced with AgeYears (see above)
Breed2 - 10,700 of almost 15,000 rows are populated with 0 (unknown), so this doesn't seem like a good feature to keep
State - initially kept this, but the correlation to AdoptionSpeed is only about 2%.  Decided to remove it
RescuerID - common sense is that this field will not have any predictive value for adoption rate
Description - this will have value in future analysis for NLP, but this will be evaluated differently in another notebook
PetID - same reason as RescuerID

```{r}
data = subset(data, select = -c(Name, Age, Breed2, State, RescuerID, Description, PetID))
```

### Show the updated data type
```{r}
sapply(data,class)
```

#Change categorical variables from integer to factors
```{r}
data <- transform(
  data,
  Type=as.factor(Type),
  Breed1=as.factor(Breed1),
  Gender=as.factor(Gender),
  Color1=as.factor(Color1),
  Color2=as.factor(Color2),
  Color3=as.factor(Color3),
  MaturitySize=as.factor(MaturitySize),
  FurLength=as.factor(FurLength),
  Vaccinated = as.factor(Vaccinated),
  Dewormed = as.factor(Dewormed),
  Sterilized = as.factor(Sterilized),
  Health = as.factor(Health),
  AdoptionSpeed = as.factor(AdoptionSpeed),
  NumColors = as.factor(NumColors)
)
```


```{r}
head(data)
```

###Output an updated csv 
```{r}
write.csv(data, './data/TrainingData/train_cleaned.csv',row.names = FALSE)
#write.csv(data, './data/TestData/test_cleaned.csv',row.names = FALSE)
```

