---
title: "DATA 621 Final Project"
author: "Group 2"
---

### Overview
There are millions of stray pets around the world, some of which are fortunate enough to be adopted while many others are not.  While adoption of a pet is often the definition of success, the rate at which a pet is adopted is also a key success factor - pets that take a long time to adopt contribute to over-crowded animal shelters and can prevent taking on new strays.  Sadly, pets that are not adopted eventually need to be euthanized.

### Learn more about the data
[About the data](https://www.kaggle.com/c/petfinder-adoption-prediction/data)

### What to Predict
Predictor (Adoption Speed) Description: Predict how quickly, if at all, a pet is adopted.

The values are determined in the following way:
0 - Pet was adopted on the same day as it was listed.
1 - Pet was adopted between 1 and 7 days (1st week) after being listed.
2 - Pet was adopted between 8 and 30 days (1st month) after being listed.
3 - Pet was adopted between 31 and 90 days (2nd & 3rd month) after being listed.
4 - No adoption after 100 days of being listed.

### This notebook...
inputs the transformation training data and creates several models using RandomForest and XGBoost 

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(randomForest)
require(xgboost)
library(Matrix)
library(caret)
library(here)
require(caTools)

#importing our functions
source(here("R","draw_confusion_matrix.R"), local = knitr::knit_global())
```

### Load Data
Load the data
```{r echo=FALSE}
# load data
data <- read.csv(here('data','TrainingData','train_cleaned.csv'))
head(data,2)
```

```{r}
data <- transform(
  data,
  Type=as.factor(Type),
  #Breed1=as.factor(Breed1),
  Gender=as.factor(Gender),
  Color1=as.factor(Color1),
  Color2=as.factor(Color2),
  Color3=as.factor(Color3),
  MaturitySize=as.factor(MaturitySize),
  FurLength=as.factor(FurLength),
  Vaccinated = as.factor(Vaccinated),
  Dewormed = as.factor(Dewormed),
  Sterilized = as.factor(Sterilized),
  Health = as.factor(Health),
  AdoptionSpeed = as.factor(AdoptionSpeed),
  NumColors = as.factor(NumColors)
)
```

### Functions
#### Function to create train and test
```{r}

source(here("R","partition_train_test_data.R"),
       local = knitr::knit_global())
```

#### Function to create RF confusion matrix
```{r}
source(here('R', 'create_rf_confusion_matrix.R'))
```

#### Function to create RF Model
```{r}
source(here('R','rf_model_func.R'))
```

#### Function to create XG Boost model
```{r}
source(here('R','xg_model_func.R'))
```

### RandomForest (RF)
Helpful article from [TowardsDataScience](https://towardsdatascience.com/random-forest-in-r-f66adf80ec9)

#### RF Model #1
```{r}
#dataframe for RF model 1
rf_model1_data = subset(data, select = c(Type,Breed1,FurLength,Gender,Sterilized,Quantity,AgeYears,Vaccinated,AdoptionSpeed))
  
rf_confusion_matrix = rf_model_func(rf_model1_data)
rf_confusion_matrix
```

```{r}
draw_confusion_matrix(rf_confusion_matrix)
```


#### RF Model #2
```{r}
#dataframe for RF model 2
rf_model2_data = subset(data, select = c(Type,AgeYears,AdoptionSpeed))
  
rf_confusion_matrix = rf_model_func(rf_model2_data)
rf_confusion_matrix
```

```{r}
draw_confusion_matrix(rf_confusion_matrix)
```

#### RF Model #3
```{r}
#dataframe for RF model 3
rf_model3_data = subset(data, select = c(Type,Breed1,FurLength,Gender,AgeYears,AdoptionSpeed))
  
rf_confusion_matrix = rf_model_func(rf_model3_data)
rf_confusion_matrix
```

### XGBoost (XG)
Helpful article from [Analytics Vidyha](https://www.analyticsvidhya.com/blog/2016/01/xgboost-algorithm-easy-steps/)
#### XG Model #1
```{r}
#dataframe for XG model 1
xg_model1_data = subset(data, select = c(Type,Breed1,FurLength,Gender,Sterilized,Quantity,AgeYears,Vaccinated,AdoptionSpeed))

xg_confusion_matrix = xg_model_func(xg_model1_data)
xg_confusion_matrix
```

#### XG Model #2
```{r}
#dataframe for XG model 2
xg_model2_data = subset(data, select = c(Type,AgeYears,AdoptionSpeed))

xg_confusion_matrix = xg_model_func(xg_model2_data)
xg_confusion_matrix
```

#### XG Model #3
```{r}
#dataframe for XG model 3
xg_model3_data = subset(data, select = c(Type,Breed1,FurLength,Gender,AgeYears,AdoptionSpeed))

#xg_confusion_matrix = xg_model_func(xg_model3_data)
#xg_confusion_matrix
```