---
title: "Data 621 Assignment 1"
subtitle: ""
author: "Group2 - Gabriella Martinez, Maliat Islam, Ken Popkin, George Cruz Deschamps, Matthew Lucich and Karim Hammoud"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: pygments
    theme: cerulean 
    toc: yes
    toc_float: yes
  pdf_document: default
  openintro::lab_report: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


In this data set we are trying to identify good and bad teams in major league baseball team’s season. We are assuming some of the predictors will be higher for good teams. We will try to predict how many times a team will win in this season.

```{r, include = FALSE}
library(tidyverse)
library(corrplot)
library(pacman)
library(here)
library(RCurl)
library(tidyverse)
library(naniar)
library(UpSetR)
library(mice)
library(dplyr)
library(Metrics)
library(MLmetrics)
```


## DATA EXPLORATION:
We can observe the response variable (TARGET_WINS) looks to be normally distributed. This supports the working theory that there are good teams and bad teams. There are also a lot of average teams.

There are also quite a few variables with missing values. and,Some variables are right skewed (TEAM_BASERUN_CS, TEAM_BASERUN_SB, etc.). This might support the good team theory. It may also introduce non-normally distributed residuals in the model. We shall see.

### Load the Data

```{r, echo=FALSE, results='asis', cache=TRUE}
train <-read.csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Assignment_1/data/moneyball-training-data.csv")

test <-read.csv("https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Assignment_1/data/moneyball-evaluation-data.csv")
```

Summary of the data
```{r}
summary(train)
summary(test)
```

Glimpse of the data
```{r}
glimpse(train)

glimpse(test)
```


```{r warning=FALSE}
train %>%
  gather(variable, value, TARGET_WINS:TEAM_FIELDING_DP) %>%
  ggplot(., aes(value)) + 
  geom_density(fill = "#3A8B63", color="#3A8B63") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```


```{r}
temp <- train %>% 
  cor(., use = "complete.obs") #%>%
  
temp[lower.tri(temp, diag=TRUE)] <- ""
temp <- temp %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  gather(Variable, Correlation, -rowname) %>%
  filter(Variable != rowname) %>%
  filter(Correlation != "") %>%
  mutate(Correlation = as.numeric(Correlation)) %>%
  rename(` Variable` = rowname) %>%
  arrange(desc(abs(Correlation))) 
```

### Correlations with Response Variable
```{r warning=FALSE}
train %>%
  gather(variable, value, -TARGET_WINS) %>%
  ggplot(., aes(value, TARGET_WINS)) + 
  geom_point(fill = "#628B3A", color="#628B3A") + 
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = "Wins")
```

```{r}
train %>% 
  cor(., use = "complete.obs") %>%
  corrplot(., method = "color", type = "upper", tl.col = "black", diag = FALSE)
```


### DATA PREPARATION


```{r}
#remove index column from both test and train data
train <- train %>% 
  select(2:17)
test <- test %>% 
  select(2:16)
```


```colSums```^[https://statisticsglobe.com/count-number-of-na-values-in-vector-and-column-in-r]

```{r}
#NA counts for the train data set
colSums(is.na(train))
```

^[https://cran.r-project.org/web/packages/naniar/vignettes/naniar-visualisation.html]
```{r}
#visulaization and percentage of NA values
vis_miss(train)
```


^[https://datavizpyr.com/visualizing-missing-data-with-barplot-in-r/]
```{r}
#alternative NA values visualization
train  %>%
  summarise_all(list(~is.na(.)))%>%
  pivot_longer(everything(),
               names_to = "variables", values_to="missing") %>%
  count(variables, missing) %>%
  ggplot(aes(y=variables,x=n,fill=missing))+
  geom_col()
```

Since 92% of the data for the TEAM_BATTING_HBP is missing, the variable has been removed from both test and train data. TEAM_BASERUN_CS is a runner up with the next highest amount of NA at 34%.

```{r}
#removes the TEAM_BATTING_HBP due to high # of NAs
train <- train %>% 
  select(-c(TEAM_BATTING_HBP))
test <- test %>% 
  select(-c(TEAM_BATTING_HBP))
```

^[https://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/R/R-Manual/R-Manual5.html]
```{r}
#creates CSV in your current working directory of R
write.csv(train, 'hw1_train_data.csv')
write.csv(test, 'hw1_test_data.csv')
```

^[https://stats.idre.ucla.edu/r/faq/how-do-i-perform-multiple-imputation-using-predictive-mean-matching-in-r/]
```{r}
#build_kmeans()

imputed_train_data <- mice(train, m=1)

imputed_test_data <- mice(test, m=1)
```

```{r}
#appends the imputed data to the original data
imp_train <- complete(imputed_train_data, "long", inc = TRUE)

imp_test <- complete(imputed_test_data, "long", inc = TRUE)
```

```{r}
#Imputation Diagnostic Checks

## labels observed data in blue and imputed data in red for y1
col <- rep(c("blue", "red")[1 + as.numeric(is.na(imputed_train_data$data$TEAM_BATTING_H))], 6)
## plots data for y1 by imputation
stripplot(TEAM_BATTING_H ~ .imp, data = imp_train, jit = TRUE, col = col, xlab = "imputation Number")
```



## BUILD MODELS

```{r echo=FALSE}
# load data
data <- read.csv('https://raw.githubusercontent.com/akarimhammoud/Data_621/main/Assignment_1/data/moneyball-training-data.csv')

cases = dim(data)[1]
features = dim(data)[2]
cat('Data for this project is', cases, 'cases and', features, 'features')
```

```{r data}
head(data,1)
```

### Create train and test data
```{r}
data$id <- 1:nrow(data)
train <- data %>% dplyr::sample_frac(.75)
test  <- dplyr::anti_join(data, train, by = 'INDEX')
```

```{r}
cases = dim(train)[1]
features = dim(train)[2]
cat('Training data for this project is', cases, 'cases and', features, 'features')
```

```{r}
cases = dim(test)[1]
features = dim(test)[2]
cat('Testing data for this project is', cases, 'cases and', features, 'features')
```

### First Model
Using a manual review, below are the features selected for the first model and the supporting reason/s.

TEAM_BATTING_H = Base hits by batters:  it's impossible to win in baseball without getting to the bases and hitting the ball is the primary means to accomplish this.

TEAM_PITCHING_H = Hits allowed: winning without a good defense is difficult and in baseball preventing the other team from getting hits is a good defense strategy.

Only two features are selected for the first model - start small and build up seems like a good approach.

<B> Create the Regression Model </B>  
```{r}
#Select the desired data for the model
rmdata <- train %>%
  select(TEAM_BATTING_H, TEAM_PITCHING_H, TARGET_WINS)

#Build the first model and produce a summary
first_model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_PITCHING_H, data = rmdata)
summary(first_model)
```

The p values are 0, which per the criteria of "keep a feature if the p-value is <0.05" recommends that we keep both these features.  But, the adjusted R-squared is TERRIBLE at around 21%.  Even though the R-squared is poor it's simple to run this model with the test data, so we'll do that next. 

```{r}
#Predict with the first model training data
first_model_predictions = predict(first_model,test)
```

```{r}
#Evaluate the first model results using RMSE
rmse(test$TARGET_WINS, first_model_predictions)
```

### Second Model
Using a manual review, below are the features selected for the second model and the supporting reason/s.

We'll keep the features from the first model (due to low p-values) and add two more features...
TEAM_FIELDING_E = Errors: errors are costly in terms of immediate impact, but could also impact the team in other ways (i.e. a high occurrence could impact team comraderie and confidence in each other)

TEAM_PITCHING_BB = Walks allowed: putting players on base for "free" is more opportunity for points 

<B> Create the Regression Model </B>  
```{r}
#Select the desired data for the model
rmdata <- train %>%
  select(TEAM_BATTING_H, TEAM_PITCHING_H, TEAM_FIELDING_E, TEAM_PITCHING_BB, TARGET_WINS)

#Build the second model and produce a summary
second_model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_PITCHING_H + TEAM_FIELDING_E + TEAM_PITCHING_BB, data = rmdata)
summary(second_model)
```

```{r}
#Predict with the second model training data
second_model_predictions = predict(second_model,test)
```

```{r}
#Evaluate the second model results using RMSE
rmse(test$TARGET_WINS, second_model_predictions)
```
The increase from two features in the first model to four features in the second model did not yield a noticeable improvement.  The Adjusted R2 on the training data improved slightly, but the RMSE for all practical purposes stayed the same at around 13; which is a poor RMSE implying that both models have poor predictive capability.

### Third Model




## SELECT MODELS

